\documentclass{letter}
\makeatletter
\let\@texttop\relax
\makeatother

\topmargin      0.0in
\headheight     0.0in
\headsep        0.0in
\oddsidemargin  0.0in
\evensidemargin 0.0in
\textheight     9.0in
\textwidth      6.5in

\begin{document}
\address{Nathanael Thompson\\
1102 Lincolnshire Drive\\
Champaign, IL 61821\\
nat@alumni.brown.edu\\
(469) 230-7125 (Mobile)}
\signature{Nathanael Thompson}
\begin{letter}{Pivotal Labs\\1701 Pearl Street\\Suite 200\\Boulder, CO 80302}

\opening{Dear Sir or Madam,}
My name is Nathanael Thompson.  I am a senior software engineer at Citrix
Systems in the Bytemobile branch located in Champaign, IL.  I have worked at
Bytemobile for the last two years since completing my PhD in Computer Science at
the University of Illinois.  While looking for software engineering jobs in the
Boulder/Denver area I discovered Pivotal Labs and I was very intrigued.

I have come to realize over the last few years that I have a strong passion for
creating beautiful software. While first lookign for jobs I was intent on
finding a position that overlapped with my research focus - mobile computing and
distributed systems.  However, I soon realized that I enjoyed more and was
interested more in the process of software construction - from designing,
authoring, testing and maintaining.  When I read about Pivotal it seemed like
a company that shares the same passions.

Technically I believe I can make significant contributions to the Pivotal team.
With a PhD in Computer Science I have a solid background in computer science
fundamentals.  My focus in disruption-tolerant networks and distributed systems
carries over into the large-scale systems and cloud platforms that Pivotal is
designing.  I have a proven ability to lead my own projects and produce results
as demonstrated by my publication record and responsibilities to lead
feature design and implementation at Bytemobile.  I have been programming in
Java since my freshman year at Brown in 1997 and spent two years working in the
Java group at Sun Microsystems (now Oracle).  I have written software to run on
hundreds of nodes in the PlanetLab research network and developed
high-performance software to run in mobile carrier networks at Bytemobile.
Although our group does not follow Agile principles, I have 
spearheaded an effort to increase unit testing as well as championing agressive
refactoring, elimination of code duplication and encapsulation.


Overall I believe I can be a successful part of the Pivotal team.
I look forward to arranging a meeting and
interacting with you further.  
I can be contacted via email at
nat@alumni.brown.edu or via cell phone at 469-230-7125.  
Thank you for your consideration.

\closing{Sincerely,}
\end{letter}
\end{document}
