\documentclass{letter}
\makeatletter
\let\@texttop\relax
\makeatother

\topmargin      0.0in
\headheight     0.0in
\headsep        0.0in
\oddsidemargin  0.0in
\evensidemargin 0.0in
\textheight     9.0in
\textwidth      6.5in

\begin{document}
\address{Nathanael Thompson\\812 Breen Drive\\Champaign, IL 61820\\nathomps@cs.uiuc.edu\\(469)
230-7125 (Mobile)}
\signature{Nathanael Thompson}
\begin{letter}{BBN Technologies\\10 Moulton Street\\Cambridge, MA 02138}

\opening{Dear Sir or Madam,}
I am a doctoral candidate in the Computer Science department at the
University of Illinois Urbana-Champaign working with Prof. Robin Kravets.  I
am expecting to receive my degree this May in the area of mobile networking.  
While working on my PhD I have developed
a desire to apply my skills as a networking and systems
researcher in a more applied setting. 
I know of BBN from the many contributions that the company has made in the
field of computer networking 
and also from my labmates, BBN interns Sam Nelson and Mehedi
Bakht and I believe BBN would be an ideal place for me to pursue
my career goals.

% Why am I good at networking
My research at the University of Illinois has focused on the area of mobile
computing and networking.  In particular, I have studied algorithms and
protocols to improve the performance of communication between end-user
devices.  My projects have included congestion control for delay-tolerant
networks, authentication management for large networks of privately-owned WiFi
access points and intelligent bandwidth aggregation across nearby wireless
routers.  Currently I am investigating location-based replication management for
delay-tolerant vehicular networks.  Working on these projects has helped me
develop an understanding of how to conduct research in computer
networking, as illustrated by my publications 
at several top mobile networking conferences.  
For several of the projects I have also created
full implementations
that have been downloaded and used by the community, demonstrating my
ability to deliver working products.  Furthermore, one project
even led to a patent application.
More broadly, I have 
developed a strong knowledge of networking and systems in general
through attendance at top conferences in the area and
graduate seminars and research classes in
distributed systems, the Internet and operating systems.  
%I also had the opportunity to deepen my knowledge of computer networking through
%my experience as a teaching assistant 
%for the graduate-level course which taught networking
%fundamentals from the MAC layer to the transport layer.
%The projects have also covered data mining and network modeling.

% I am good at networking
As a networking and systems researcher at the University of Illinois I
have been able to develop a strong understanding of computer networks and the
Internet.
I have had the opportunity to work
on several networking projects including the large-scale authentication system
AGE, the network monitoring tool NLANR Network Advisor at NCSA, and my current
research in delay-tolerant networks of several hundred devices.
In addition, I have also
been able to develop a strong knowledge of networking and systems in general
through attendance at top conferences and
graduate-level seminars and research classes in the area.
As a teaching assistant
for the graduate-level networking course 
I further deepened my knowledge of computer networking
by teaching networking
fundamentals from the MAC layer to the transport layer.
I have demonstrated an ability to apply my knowledge to generate high quality
work in a timely manner by publishing my research at several top mobile
networking conferences.  For several of my projects I have also created
full implementations
which have been downloaded and used by the community.  Furthermore, one project
even led to a patent application.
%The projects have also covered data mining and network modeling.
I was able to
develop an extensive familiarity with the Java APIs and technologies
and grow a strong foundation in object-oriented programming
as an engineer at Sun Microsystems.% processing Java bug reports. 



% I am a good computer scientist
Not only has my time as a graduate student taught me how to effectively
understand and evaluate large-scale systems and generate new ideas,
it has also taught me how to effectively communicate those ideas.  
By working in collaborative research projects between research groups and
different organizations I have learned how to clearly express my thoughts to 
a variety of people and synthesize my and their proposals in
order to come to a shared solution.
Preparing conference and poster presentations and writing publications has
taught me how to clearly articulate my thoughts in a concise manner.
Also, overseeing younger graduate students and working as a teaching assistant
for
both graduate and undergraduate classes forced me to learn how to communicate
with people with different backgrounds and experience in my area.
Finally, my previous experience in industry working at Sun Microsystems and
NCSA has trained me to work within a team in order to deliver a product.


% closing
I am sure that these skills will help me to be an effective part of BBN,
continuing BBN's tradition of pioneering work in the technology fields.
I look forward to arranging a meeting and interacting with you further.  
%I can be contacted via email at
%nathomps@cs.uiuc.edu or via cell phone at 469-230-7125.  
Thank you for your consideration.

\closing{Sincerely,}
\end{letter}
\end{document}
